Unsupervised training of maximum-entropy models for lexical selection in rule-based machine translation

### 2015-02-11 2015-eamt-melsa/2015-eamt-melsa.tex
8 pages \o/
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@58848 72bbbca6-d526-0410-a7d9-f06f51895060

\subsection{Prior work}
%@@@ \comment{This section should be ironed out a bit.}  
In their
pioneering work, \cite{dagan1994word} used the term
\emph{word sense disambiguation} to refer to what is actually target
word selection in machine translation; they used a parser to identify
some syntactic relations such as subject--object or subject--verb.
After generating all the possible translations for a given
input sentence using an ambiguous bilingual dictionary, they extract the syntactic
tuples from the target language and count the frequency in a previously-trained target-language 
model of tuples.  They use maximum-likelihood estimation to calculate the probability of a given 
target language tuple being the translation of a given source language tuple, with
an automatically determined threshold for confidence.

Then \cite{berger1996} illustrated the use of maximum-entropy classifiers
on the specific problem of lexical selection in IBM-style word-based
statistical machine translation. Other authors
\citep{melero07a} have used target-language models to rank %@@@ carbonell06
the translations resulting from all possible combinations of lexical
selections.  Nowadays, in state-of-the-art phrase-based statistical
machine translation \citep{koehn2009statistical}, lexical selection is
taken care of by a combination of the translation model and the
language model. The translation model provides probabilities of
translation between words or word sequences (often referred to as
\emph{phrases}) in the source and target language. The target-language
model provides probabilities of word sequences in the target language.
As regards rule-based machine translation, \cite{sanchez07a} proposed 
to use information from the target language to train a statistical 
lexical-selection module but did not provide any experiments. 
%@@@ \todo[inline]{FMT: integrate \citep{sanchez07a} somewhere here}
\cite{marechek10} trained a maximum-entropy lexical selector for their TectoMT rule-based system
using a bilingual corpus.  More recently, \cite{tyers12a} presented a
method of lexical selection for rule-based machine translation based
on rules which select or remove translations in fixed-length contexts,
along with a training method for learning the rules from a
word-aligned parallel corpus.

@@

There are a number of approaches to creating evaluation corpora for
lexical selection in the literature. \cite{vickrey05} use a parallel
corpus to make annotated test and training sets for experiments in
lexical selection applied to a simplified translation problem in
SMT. They use word alignments from {\sc giza++} \citep{och03a} to
annotate source language words with their translations from the
reference translation in the parallel corpus. One disadvantage of this
method is that only one translation is annotated per source language
word, meaning that accuracies may be lower because of missing
translations --- this happens when the system chooses a translation
which is adequate, but is not found in the reference translation. A
second disadvantage is that the word alignments may not be 100\%
reliable, which decreases the accuracy of the annotated corpus. An
alternative method is described by \cite{zinovjeva2000}, who manually
tags a selection of ambiguous words in English sentences with their
translation in Swedish. %The size of the annotated corpus is not
%mentioned. 
%\cite{specia2005a} use a parallel corpus, but eschew using
%word alignment in favour of using a bilingual dictionary and heuristic
%rules based on part-of-speech and relative position. The corpus was
%reviewed manually before being used in further experiments
%\citep{specia2005}.

@@

The ideal situation would be to have, as described by
\cite{zinovjeva2000}, a hand-annotated evaluation corpus for testing
the performance of the lexical-selection module. That is, the output
of the lexical transfer module which had been disambiguated by hand by
one or more human annotators. As this did not exist for any language
pair, we decided to automatically annotate a test set for each
language pair using a process similar to that described by
\cite{vickrey05}.

@@

After this process, we selected 1,000 sentence pairs at random for
testing (\texttt{test}), 1,000 for development (\texttt{dev})\footnote{The development corpus  
  was used for checking the value for frequency pruning of features.} and 
left the remainder for training. This is a smaller
number of sentences than used in typical evaluations (such as the WMT
series, \cite{callisonburch2012}), but was motivated by the fact that
the smallest corpus (for Breton--French) after sentence discarding was
only little over 4,000 sentences long. Table~\ref{table:input-corp}
gives statistics about the size of the input corpora, and how many
sentences were left after processing for testing, training and
development. %Table~\ref{table:test-corp} gives information about the
%test corpora.

@@

\begin{itemize}
\addtolength{\itemsep}{-0.05in}
\item \textbf{Linguist-chosen defaults}. 
%! \comment{FSM: This should go to the section explaining Apertium.}  
  A bilingual dictionary in an
  Apertium language pair contains correspondences between lexical
  forms. The dictionaries allow many lexical forms to translate to one
  lexical form. For example, \emph{ahizpa} and \emph{arreba} in Basque
  both translate to \emph{hermana} `sister' in Spanish.\footnote{The
    difference being that \emph{ahizpa} is a sister of a female, and
    \emph{arreba} is a sister of a male.} But a single lexical form
  may not have more than one translation without further
  processing. If there are many possible translations of a lexical
  form, then one must be marked as the \emph{default} translation. For
  example, \emph{pisu} in Basque translates to \emph{peso} `weight'
  and \emph{piso} `flat' in Spanish. If the translation correspondence
  \emph{pisu $\rightarrow$ peso} is perceived as the most frequent or
  the most general, it is marked as the default by the author of the dictionary. % linguist who is
%  creating the bilingual dictionary.

>>>>

\subsection{Prior work}
%@@@ \comment{This section should be ironed out a bit.}  
%In their pioneering work, 
\cite{dagan1994word} used the term
\emph{word sense disambiguation} to refer to what is actually target
word selection in machine translation; they used a parser to identify
some syntactic relations such as subject--object or subject--verb.
After generating all the possible translations for a given
input sentence using an ambiguous bilingual dictionary, they extract the syntactic
tuples from the target language and count the frequency in a previously-trained target-language 
model of tuples.  They use maximum-likelihood estimation to calculate the probability that a given 
target language tuple is the translation of a given source language tuple, with
an automatically determined confidence threshold.

Then \cite{berger1996} illustrated the use of maximum-entropy classifiers
on the specific problem of lexical selection in IBM-style word-based
statistical machine translation. Other authors
\citep{melero07a} have used target-language models to rank %@@@ carbonell06
the translations resulting from all possible combinations of lexical
selections.  Nowadays, in state-of-the-art phrase-based statistical
machine translation \citep{koehn2009statistical}, lexical selection is
taken care of by a combination of the translation model and the
language model. The translation model provides probabilities of
translation between words or word sequences (often referred to as
\emph{phrases}) in the source and target language. The target-language
model provides probabilities of word sequences in the target language.
%As regards rule-based machine translation, \cite{sanchez07a} proposed 
%to use information from the target language to train a statistical 
%lexical-selection module but did not provide any experiments. 
%@@@ \todo[inline]{FMT: integrate \citep{sanchez07a} somewhere here}
\cite{marechek10} trained a maximum-entropy lexical selector for their TectoMT rule-based system
using a bilingual corpus.  More recently, \cite{tyers12a} presented a
method of lexical selection for rule-based machine translation based
on rules which select or remove translations in fixed-length contexts,
along with a training method for learning the rules from a
word-aligned parallel corpus.

@@

There are a number of approaches to creating evaluation corpora for
lexical selection in the literature. \cite{vickrey05} use a parallel
corpus to make annotated test and training sets for experiments in
lexical selection applied to a simplified translation problem in
SMT. They use word alignments from {\sc giza++} \citep{och03a} to
annotate source language words with their translations from the
reference translation in the parallel corpus. One disadvantage of this
method is that only one translation is annotated per source language
word, meaning that accuracies may be lower because of missing
translations --- this happens when the system chooses a translation
which is adequate, but is not found in the reference translation. A
second disadvantage is that the word alignments may not be 100\%
reliable, which decreases the accuracy of the annotated corpus. An
alternative method is described by \cite{zinovjeva2000}, who manually
tags ambiguous words in English sentences with their % a selection of
translation in Swedish. %The size of the annotated corpus is not
%mentioned. 
%\cite{specia2005a} use a parallel corpus, but eschew using
%word alignment in favour of using a bilingual dictionary and heuristic
%rules based on part-of-speech and relative position. The corpus was
%reviewed manually before being used in further experiments
%\citep{specia2005}.

@@

%The ideal situation would be to have, as described by
%\cite{zinovjeva2000}, a hand-annotated evaluation corpus for testing
%the performance of the lexical-selection module. That is, the output
%of the lexical transfer module which had been disambiguated by hand by
%one or more human annotators. As this did not exist for any language
%pair, we decided to automatically annotate a test set for each
%language pair using a process similar to that described by
%\cite{vickrey05}. 
%
Ideally we would have had a hand-annotated evaluation corpus, as
described by \cite{zinovjeva2000}, but as this did not exist, we 
decided to automatically annotate a test set using a process similar to 
that described by \cite{vickrey05}.

@@

After this process, we selected 1,000 sentence pairs at random for
testing (\texttt{test}), 1,000 for development (\texttt{dev})\footnote{The development corpus  
  was used for checking the value for frequency pruning of features.} and 
left the remainder for training. 
%This is a smaller
%number of sentences than used in typical evaluations (such as the WMT
%series, \cite{callisonburch2012}), but was motivated by the fact that
%the smallest corpus (for Breton--French) after sentence discarding was
%only little over 4,000 sentences long. 
Table~\ref{table:input-corp}
gives statistics about the size of the input corpora, and how many
sentences were left after processing for testing, training and
development. %Table~\ref{table:test-corp} gives information about the
%test corpora.

@@

\begin{itemize}
\addtolength{\itemsep}{-0.05in}
\item \textbf{Linguist-chosen defaults}. 
%! \comment{FSM: This should go to the section explaining Apertium.}  
  A bilingual dictionary in an
  Apertium language pair contains correspondences between lexical
  forms. The dictionaries allow many lexical forms to translate to one
  lexical form. 
%For example, \emph{ahizpa} and \emph{arreba} in Basque
%  both translate to \emph{hermana} `sister' in Spanish.\footnote{The
%    difference being that \emph{ahizpa} is a sister of a female, and
%    \emph{arreba} is a sister of a male.} 
  But a single lexical form
  may not have more than one translation without further
  processing. If there are many possible translations of a lexical
  form, then one must be marked as the \emph{default} translation. 
%For
%  example, \emph{pisu} in Basque translates to \emph{peso} `weight'
%  and \emph{piso} `flat' in Spanish. If the translation correspondence
%  \emph{pisu $\rightarrow$ peso} is perceived as the most frequent or
%  the most general, it is marked as the default by the author of the dictionary. % linguist who is
%  creating the bilingual dictionary.
